---
title: "1536_test_plate_plan"
author: "Jamie Auxillos"
date: "01/11/2020"
output: html_document
---


```{r setup,warning=FALSE,message=FALSE,echo=FALSE}
## knitr options for report generation
knitr::opts_chunk$set(warning=FALSE,message=FALSE,echo=TRUE,cache=FALSE,
                      results="show",
                      fig.path="figure_1536_pilot")


library(tidyverse)
library(cowplot)
library(tidyqpcr)
library(here)

 # set default theme for graphics
theme_set(theme_cowplot(font_size=11) %+replace% 
              theme(panel.border=element_rect(colour = "grey50",
                                            linetype = "solid",size=0.5),
                    strip.background = element_blank()))

```

## Plan for reaction setup

Previously Edward optimised the 1536 qPCR method using a 1 ul total reaction volume. This 1 ul reaction consist of:
  - 0.8uL SYBR Green 2X master mix, diluted 1.6-fold ( 5 mix : 3 H20 )
  - 0.1uL of 4uM primer pair
  - 0.1uL of RT reaction, diluted 1:4

Primers and cDNA are loaded on a 384PP (15ul-65 ul) plate and dispensed into the 1536 plate using the Echo. Then the sybr green is loaded into the 1536 plate using the Certus bulk dispenser.

mCherry primer mix
  - 0.1 ul of 4 uM primer pair
  - Therefore for 1536 samples, 153.6 ul is needed (1536 x 0.1 = 153.6)
  - 50 ul of the primer mix is loaded in the source plate

cDNA
  - For +RT wells
        - There are 1024x +RT wells, half (512x) are with pPGK1-mCh and the other half (512x) with POT1-ccdB
        - So the +RT cDNA is diluted 1:4 and 56.2 ul of each cDNA is loaded in the source plate
  - For -RT wells
        - There are 512x -RT wells, half (256x) are with pPGK1-mCh and the other half (256x) with POT1-ccdB
        - So the -RT cDNA is diluted 1:4 and 27.6 ul of each cDNA is loaded in the source plate


Sybr green dilution
  - Dilute the Agilent 2x sybr green 5:3 (MasterMix:water)
        - The amount of the diluted sybr green  used in the final reaction is 0.8 ul
        - Therefore, we will need (1536 x 0.8 = 1228.8) 1300 ul of the diluted 5:3 sybr green
        - for this, 812.5 ul of sybr green is diluted in 487.5 ul of ddH2O
  - Diluted sybr green is loaded using the bulk dispenser


## Label and plan plates

```{r label_destinationplate,dependson="plate_functions"}

# Names of target primers
Names <- c("mCh-7") 

# Repeats of gene names to account for resting multiple primer sets
probes <- c(rep(Names,times=32))

# Labelss for strains
strain_names <- c("pPGK1-mCh", "POT")
strain_times <- c(rep(strain_names,times=16))

# Rowkey for destination plate
rowkey <- data.frame(
                     well_row = make_row_names_lc1536(),
                     strain= strain_times,
                     target_id=probes)

# Names for prep_type
condition <- c("+RT","-RT", "+RT")

# Colkey for destination plate
colkey <- data.frame(well_col = 1:48,
                     prep_type=c(rep(condition, times=16)))

qpcr_plateplan <- label_plate_rowcol(
  plate=create_blank_plate(well_row = make_row_names_lc1536(),
                              well_col = 1:48),
                              rowkey=rowkey, colkey=colkey) %>%
  mutate(sample_id = paste(strain, prep_type,sep="_"))

```


## Display the plate plan

```{r display_plates,fig.height=25,fig.width=35}
display_plate(qpcr_plateplan) 
```


```{r label_sourceplate,dependson="plate_functions"}

# Names of primers
Names <- c("mCh-7") 

# Repeats of gene names to account for resting multiple primer sets
probes <- c(rep(Names,times=4))

# Primer rowkey
primer_rowkey <- data.frame(
                     well_row = LETTERS[1:4],
                     target_id=probes)
# Primer colkey
primer_colkey <- data.frame(well_col = 3:4)

# Primer plate
primer_plateplan <- label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:4],
                              well_col = 3:4),
                              rowkey=primer_rowkey, colkey=primer_colkey)




# Names of cDNA
strain_names <- c("pPGK1-mCh", "POT")
strain_times <- c(rep(strain_names,times=4))
cdna_names <- c("+RT", "+RT", "+RT", "-RT")
cdna_times <- c(rep(cdna_names,each=2))

# cDNA rowkey
cdna_rowkey <- data.frame(
                     well_row = LETTERS[1:8],
                     strain=strain_times,
                     prep_type=cdna_times)
# cDNA colkey
cdna_colkey <- data.frame(well_col = 1)

# cDNA plate
cdna_plateplan <- label_plate_rowcol(
  plate=create_blank_plate(well_row = LETTERS[1:8],
                              well_col = 1),
                              rowkey=cdna_rowkey, colkey=cdna_colkey)   %>%
  mutate(sample_id = paste(strain, prep_type,sep="_"))  

```

```{r pick_list_functions}
pair_qpcr_wells_to_source_wells_by_liquid_type <- function(qpcr_plateplan, source_plateplan, liquid_type = "sample"){
  
  # check liquid_type is either sample or target
  if(!liquid_type %in% c("sample", "target")) stop("Incorrect liquid type.")
  
  # combine prep_type and sample_id for both plates
  if(liquid_type == "sample"){
    source_plateplan <- unite(source_plateplan,
                              sample_prep,
                              sample_id,
                              prep_type,
                              remove = FALSE)
    
    qpcr_plateplan <- unite(qpcr_plateplan,
                            sample_prep,
                            sample_id,
                            prep_type,
                            remove = FALSE)
  }
  
  # check that the total volume of each liquid across all wells in the source plate is 
  # greater than the required amount across all wells in the qpcr_plate
  check_source_plate_volume(source_plateplan, qpcr_plateplan, liquid_type)
  
  # Ensure all target_id and sample_id wells in source/qpcr plates contain the same total volumes, else automatic assignment is too difficult
  if(length(unique(source_plateplan %>%
                   filter(!is.na(.data[[paste0(liquid_type, "_id")]])) %>%
                   pull(.data$max_well_volume))) > 1) {
    stop("Cannot automatically assign wells if max liquid volumes of any ", liquid_type, " varies across different wells of the source plate") }
  
  if(length(unique(qpcr_plateplan %>%
                   filter(!is.na(.data[[paste0(liquid_type, "_id")]])) %>%
                   pull(.data[[paste0(liquid_type, "_volume")]]))) > 1) {
    stop("Cannot automatically assign wells if requested liquid volumes of any ", liquid_type, " varies across different wells of the qpcr plate")
                   }
    
    # Calculate how many qpcr wells can be supplied from a single source well
  qpcr_wells_per_source_well <- floor(unique(source_plateplan %>%
                                         filter(!is.na(.data[[paste0(liquid_type,
                                                                     "_id")]])) %>%
                                           mutate(well_volume = 
                                                    max_well_volume - min_well_volume) %>%
                                         pull(.data$well_volume)) / 
                                         unique(qpcr_plateplan %>%
                                         filter(!is.na(.data[[paste0(liquid_type,
                                                                     "_id")]])) %>%
                                         pull(.data[[paste0(liquid_type, "_volume")]])))
    
    
    # Duplicate source wells to equal number of qpcr wells that can be supplied
  total_avaliable_aliquots <- source_plateplan %>%
    filter(!is.na(.data[[paste0(liquid_type,"_id")]])) %>%
    uncount(qpcr_wells_per_source_well)
  
  # pair wells
  if(liquid_type == "sample"){
  
    paired_qpcr_plateplan <- qpcr_plateplan %>%
      filter(!is.na(.data[["sample_id"]])) %>%
      group_by(.data[["sample_prep"]]) %>%
      mutate(source_well = assign_source_wells(.data[["sample_prep"]],
                                               "sample_prep",
                                               total_avaliable_aliquots)) %>%
      ungroup() %>%
      select(-sample_prep)
  }
  else{
    paired_qpcr_plateplan <- qpcr_plateplan %>%
      filter(!is.na(.data[[paste0(liquid_type,"_id")]])) %>%
      group_by(.data[[paste0(liquid_type,"_id")]]) %>%
      mutate(source_well = assign_source_wells(.data[[paste0(liquid_type,"_id")]],
                                               paste0(liquid_type,"_id"),
                                               total_avaliable_aliquots))
  }
  
  paired_qpcr_plateplan %>%
    rename_with((function(col_names)paste0(liquid_type, "_source_well")),source_well)
  
}

check_source_plate_volume <- function(source_plateplan, qpcr_plateplan, liquid_type = "sample") {
  # check liquid_type is either sample or target
  if(!liquid_type %in% c("sample", "target")) stop("Incorrect liquid type.")
  
  # Select column to group by
  if(liquid_type == "sample") grouping_variable = "sample_prep"
  else grouping_variable = paste0(liquid_type, "_id")
  
  total_required_volume <- qpcr_plateplan %>%
    group_by(.data[[grouping_variable]]) %>%
    summarise(total_required_volume = sum(.data[[paste0(liquid_type, "_volume")]])) %>%
    pull(total_required_volume)
  
  total_available_volume <- source_plateplan %>%
    filter(!is.na(.data[[paste0(liquid_type, "_id")]])) %>%
    group_by(.data[[grouping_variable]]) %>%
    summarise(total_available_volume = sum(max_well_volume - min_well_volume)) %>%
    pull(total_available_volume)
  
  if (sum(total_required_volume > total_available_volume) != 0) {
    stop(
      paste0("Source plate does not have enough total ", 
             paste0(qpcr_plateplan %>% 
                      arrange(.data[[grouping_variable]]) %>% 
                      distinct(.data[[grouping_variable]]) %>% 
                      filter(total_required_volume > total_available_volume) %>% 
                      pull(.data[[grouping_variable]]), collapse = " & "),
             " to fill qpcr plate."))
  }
}

assign_source_wells <- function(liquid_id, liquid_column, total_avaliable_aliquots){
  paired_source_wells <- total_avaliable_aliquots %>%
    filter(.data[[liquid_column]] == liquid_id[[1]]) %>%
    pull(well)
  
  if(length(paired_source_wells) < length(liquid_id)) stop("Automatic source-qpcr well assignment failed, you may have to manually assign wells")
  
  paired_source_wells[1:length(liquid_id)]
}

generate_picklist <- function(qpcr_plate_well_map, file){
  pick_targets <- qpcr_plate_well_map %>%
    filter(!is.na(sample_id)) %>%
    transmute(Source = target_source_well,
              Target = well,
              Volume = target_volume,
              LType = "Target",
              LID   = target_id)

pick_samples <- qpcr_plate_well_map %>%
    filter(!is.na(sample_id)) %>%
    transmute(Source = sample_source_well,
              Target = well,
              Volume = sample_volume,
              LType = "Sample",
              LID   = sample_id)

picklist <- bind_rows(pick_targets, pick_samples) %>%
  separate(Source, into = c("Source_row", "Source_col"), remove = FALSE, sep = "(?<=[A-Za-z])(?=[1-9])") %>%
  separate(Target, into = c("Target_row", "Target_col"), remove = FALSE, sep = "(?<=[A-Za-z])(?=[1-9])") %>%
  arrange(Source_row, as.integer(Source_col), Target_row, as.integer(Target_col)) %>%
  select(-Source_row, -Source_col, -Target_row, -Target_col)

cat(
"# qPCR-picklist-2020-11-26-Echo-1536_testplate.txt
# Picklist for 1536-well qPCR loading with Echo liquid handler
# Four cDNA samples (2x +RT and 2x -RT) used with 1 target_id (primer)
# Sam Haynes and Jamie Auxillos, 26 Nov 2020
#
# Plates:
#  Source: 384PP 2.0, Labcyte PP-0200
#  Target: 1536-well PCR plate
# 
# Columns:
#   Source: Well in source/loading plate
#   Target: Well in target/qPCR plate
#   Volume: Liquid volume to transfer in nl
#   LType:  Type of liquid being loaded
#   LID:    Sample ID/name of liquid being loaded
# 
",
    file=file,sep="")

write_tsv(picklist,
          file,
          append=TRUE,col_names=TRUE)

}
```

```{r create_picklist}

# Add the max/min volumes of liquid in each well. The primer and cdna plates can have different volumes to each other but must be the same across each individual plate.

primer_plateplan <- primer_plateplan %>%
  mutate(max_well_volume = 50, min_well_volume = 20)

cdna_plateplan <- cdna_plateplan %>%
  mutate(max_well_volume = 50, min_well_volume = 20)

source_plateplan <- bind_rows(primer_plateplan, cdna_plateplan)

# add primer and cdna volumes to each qpcr well (must be the same value across all wells but the target and sample volumes can differ from each other)
 
qpcr_plateplan <- qpcr_plateplan %>%
    mutate(target_volume = 0.1, sample_volume = 0.1)

# Map source and qpcr plate wells

qpcr_plate_well_map <- pair_qpcr_wells_to_source_wells_by_liquid_type(qpcr_plateplan, source_plateplan, liquid_type = "target") %>%
  pair_qpcr_wells_to_source_wells_by_liquid_type(source_plateplan, liquid_type = "sample")

# Create pick list

generate_picklist(qpcr_plate_well_map, here("qPCR-picklist-2020-11-26-Echo-1536_testplate.txt"))
```
